---
title: "MA 677 Final Project"
author: "Shuoqi Huang"
date: "5/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, readxl,ggplot2,dplyr, BayesFactor)
```

## Introduction      

The goal of this project is to test the idea that a state that has more airports or enplanements has a significant difference in confirmed cases than a state that has a relatively low capacity. The reason why I choose to do this project is that large international airports perhaps are potentially helpful for spreading the Coronavirus. For example, Hartsfield - Jackson Atlanta International (ATL), Los Angeles International (LAX), and Chicago O'Hare International (ORD) are the three largest airports by checking the enplanements. Therefore, I decided to check this idea by doing Bayesian hypothesis testing. 

## Data        

I get the state confirmed cases data from Havard dataverse (https:// dataverse.harvard.edu/dataverse/2019ncov). Moreover, I explore data of Public-use airports and the number of enplanements from the Federal Aviation Administration.         

```{r load data,echo = FALSE, message= F}
# Read data 

#confirmed cases by states 
state = read_csv("data_1.csv") 

names(state)[names(state) == "NAME"] <- "ST"
names(state)[names(state) == "5/5/20"] <- "Confirmed cases"
# Number of airports by states
airport = read_excel("airports.xlsx")

f_airport = airport %>% filter( ST %in% c( "AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID","IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN","MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY","NC", "ND", "OH", "OK", "OR", "PA", "RI","SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY", "DC")  )

# enplanements by states 
enplanement = read_excel("enplanements.xlsx")

f_enplanement = enplanement %>% filter( ST %in% c( "AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID","IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN","MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY","NC", "ND", "OH", "OK", "OR", "PA", "RI","SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY", "DC")  )

```



| State  | Number of Airports | State | Number of Airports | State | Number of Airports |
|--------|--------------------|-------|--------------------|-------|--------------------|
| AK     | 406                | KY    | 60                 | NY    | 140                |
| AL     | 92                 | LA    | 74                 | OH    | 170                |
| AR     | 99                 | MA    | 40                 | OK    | 139                |
| AZ     | 79                 | MD    | 47                 | OR    | 97                 |
| CA     | 254                | ME    | 71                 | PA    | 130                |
| CO     | 76                 | MI    | 231                | RI    | 8                  |
| CT     | 23                 | MN    | 153                | SC    | 68                 |
| DC     | 4                  | MO    | 129                | SD    | 73                 |
| DE     | 11                 | MS    | 80                 | TN    | 80                 |
| FL     | 128                | MT    | 128                | TX    | 394                |
| GA     | 109                | NC    | 114                | UT    | 46                 |
| HI     | 14                 | ND    | 89                 | VA    | 65                 |
| IA     | 121                | NE    | 85                 | VT    | 16                 |
| ID     | 123                | NH    | 25                 | WA    | 138                |
| IL     | 110                | NJ    | 45                 | WI    | 133                |
| IN     | 115                | NM    | 63                 | WV    | 36                 |
| KS     | 138                | NV    | 49                 | WY    | 41                 |



We can see that Alaska has the largest number of airports, but the latest confirmed cases are 369 on May 5th. Alabama has 8437 confirmed cases with only 92 airports. Therefore, I decided to divide states into two groups by checking the enplanements.     


\newpage
The FAA only provides enplanements by each airport in some states. Therefore, I calculate the total enplanements by states and divide them into two groups by setting the checkpoint enplanes equal to 10,000,000.   


|    | ST | Capacity  | Class |    | ST | Capacity  | Class |
|----|----|-----------|-------|----|----|-----------|-------|
| 1  | CA | 117968341 | High  | 25 | LA | 7795240   | Low   |
| 2  | FL | 91150567  | High  | 26 | KY | 6844603   | Low   |
| 3  | TX | 86428773  | High  | 27 | IN | 5623344   | Low   |
| 4  | NY | 54592785  | High  | 28 | WI | 5475554   | Low   |
| 5  | GA | 53712314  | High  | 29 | SC | 5227513   | Low   |
| 6  | IL | 51892058  | High  | 30 | AK | 4985422   | Low   |
| 7  | CO | 33377605  | High  | 31 | OK | 3653023   | Low   |
| 8  | NC | 31217947  | High  | 32 | CT | 3317026   | Low   |
| 9  | VA | 27857139  | High  | 33 | NM | 2858986   | Low   |
| 10 | WA | 26956708  | High  | 34 | NE | 2709190   | Low   |
| 11 | NV | 26067981  | High  | 35 | AL | 2568618   | Low   |
| 12 | AZ | 24762641  | High  | 36 | ID | 2348824   | Low   |
| 13 | NJ | 23770909  | High  | 37 | MT | 2193321   | Low   |
| 14 | PA | 21681560  | High  | 38 | RI | 2151912   | Low   |
| 15 | MI | 20345855  | High  | 39 | IA | 2082586   | Low   |
| 16 | MA | 20314174  | High  | 40 | AR | 1934533   | Low   |
| 17 | MN | 18795176  | High  | 41 | ME | 1441318   | Low   |
| 18 | HI | 17792964  | High  | 42 | ND | 1095631   | Low   |
| 19 | MO | 14128339  | High  | 43 | NH | 1014421   | Low   |
| 20 | MD | 13459863  | High  | 44 | KS | 953132    | Low   |
| 21 | UT | 12488751  | High  | 45 | MS | 943091    | Low   |
| 22 | TN | 12010620  | High  | 46 | SD | 903958    | Low   |
| 23 | OR | 11313244  | High  | 47 | VT | 664433    | Low   |
| 24 | OH | 10266738  | High  | 48 | WY | 603422    | Low   |
|    |    |           |       | 49 | WV | 374922    | Low   |



```{r, echo= F}
## Dara wrangling 
# calculate all enplanements by each states 
c_enplanement = f_enplanement %>% group_by(ST) %>% mutate(total = sum(`CY 18 Enplanements`))%>% distinct()
ht_data = c_enplanement %>% select(c("ST", "total")) %>% distinct()
ht_data = arrange(ht_data, desc(total))

# Check the mean of enplanements
#mean(ht_data$total)

# Create classes
ht_data$Class = ifelse(ht_data$total > 10000000, "High" , "Low")

# save csv 
#write.csv(ht_data, "ht_data.csv")

# Final data for testing 
data = inner_join(ht_data,state, by = "ST") %>% select( -c('fips'))

```

## Method    

Unlike the null hypothesis testing, the Bayesian hypothesis testing is really simple to understand. I set the following hypothesis:    

$H_{0}:$ two samples have no difference that implies transportation does not affect on confirmed cases   

$H_{1}:$ two samples have differences that implies transportation  affects on confirmed cases   


The Bayes factor plays a similar role to the p-value in the traditional hypothesis testing. I will use the Bayes factor to interpret my result. 

## Result 
```{r, echo= F}
#factor class
data$Class = factor(data$Class)
data$logConfirmedCase = log(data$`Confirmed cases`)
```

First, we plot the distribution of these two groups. We can see that these two classes have different distributions. 
```{r, echo = F}
# Distribution plot 
p<-ggplot(data, aes(x=`Confirmed cases`, fill= Class)) +
  geom_density(alpha=0.4)
p

```

We can see from the previous plot that the data is right-skewed, so I perform a logarithmic transformation. Here is a summary of the data. The number of confirmed cases appears to be affected by transportation. 

```{r, echo = F}
#Boxplot 
plot( logConfirmedCase ~ Class, data = data, main = "Confirmed Cases")
```


Before doing the Bayesian factor, I use a traditional t-test to show the result. 
```{r, echo= F}
t.test(logConfirmedCase~ Class, data = data, var.eq=TRUE)
```

The p-value is really small so that I have strong evidence to reject the null hypothesis.

Let's now find the value of the Bayes factor. 

```{r,echo= F, warning= F}
# Bayes factor using BayesFactor package
bf = ttestBF( formula = logConfirmedCase ~ Class, data = data)
bf
```

The Bayes factor is 610.2352.     

\newpage
## Conclusion 

I will use the Kass and Raftery (1995) table to interpret the result. According to Kass and Raftery (1995):

| Bayes factor | Interpretation       |
|--------------|----------------------|
| 1 - 3        | Negligible evidence  |
| 3 - 20       | Positive evidence    |
| 20 - 150     | Strong evidence      |
| >150         | Very strong evidence |


Therefore, the data are 610.2352 times as probable under the alternative as under the null hypothesis. There is very strong evidence that transportation affects the number of confirmed cases. 



